The \dealii{} Library, Version 9.2

### 2020-06-01 9.2/paper.tex
Add step-70 description.

\hypersetup{
  pdfauthor={
     Daniel Arndt,
     Wolfgang Bangerth,
     Bruno Blais,
     Thomas C. Clevenger,
%     Denis Davydov,
    Marc Fehling,
%     Daniel Garcia-Sanchez,
    Alexander V. Grayver,
%     Graham Harper,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Peter Munch,
%     Ross Maguire Kynch,
    Matthias Maier,
    Jean-Paul Pelteret,
    Reza Rastak,
    Bruno Turcksin,
    David Wells
    Zhuoran Wang,
  },
  pdftitle={The deal.II Library, Version 9.2, 2020},
}

@@

\author[*1]{Bruno~Turcksin}
%
\author[14]{David Wells}
\affil[14]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

\author[15]{Zhuoran Wang}
\affil[15]{Department of Mathematics, Colorado State University, Fort Collins,
          CO 80523-1874, USA.
{\texttt{zhrwang@math.colostate.edu}}}
\renewcommand{\labelitemi}{--}

@@

The major changes of this release are:
%
\begin{itemize}
\item xy \todo[inline]{Update once we have the subsections in Section 2;
  provide cross-references to each of these subsections}
%   \item Improved support for automatic differentiation (see
%     Section~\ref{subsec:ad}),
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item \dealii{} had decent support for solving complex-valued problems
  for a while already
  (e.g., ones in quantum mechanics -- like the equation used in the
  \texttt{step-58} tutorial program covered below -- or for
  time-harmonic problems). However, there were two
  areas in which support was missing. First, the UMFPACK direct solver
  packaged with \dealii{} did not support solving complex-valued
  linear problems. This has now been addressed: UMFPACK actually can
  solve such systems, we just needed to write the appropriate
  interfaces. Second, the \texttt{DataOut} class that is responsible
  for converting nodal data into information that can then be written
  into files for visualization did not know how to deal with vector-
  and tensor-valued fields whose components are complex numbers. An
  example for this is to solve the time-harmonic version of the
  Maxwell equations that has the electric and magnetic fields as
  solution. This, too, has been addressed in this release.
\item The new \texttt{DiscreteTime} class provides a more
  consistent, more readable, and less error-prone approach to control
  time-stepping algorithms within time-dependent simulations.
  The interface of this class is designed to be minimal
  to enforce a number of important programming invariants, reducing
  the possibility of mistakes in the user code.
  An example is that \texttt{DiscreteTime} ensures that the final time step ends
  precisely on a predefined end time, automatically
  lengthening or shortening to final time step.
\item A key component of \dealii{} are the \texttt{FEValues} and
  \texttt{FEFaceValues} classes that evaluate finite element functions
  at quadrature points located on cells and faces of a cell,
  respectively~\cite{BangerthHartmannKanschat2007}. This release now
  contains a class \texttt{FEInterfaceValues} that considers the
  restriction of the shape functions from \textit{both} sides of a face and
  allows evaluating jumps and averages of shape functions along this
  face. These are common components of the bilinear forms of
  discontinuous Galerkin schemes (as well as schemes for fourth-order
  equations, see the discussion of \texttt{step-47} below) and greatly
  simplify the implementation of these methods.
\item Previously, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer}
  class, which transfers local quadrature point data onto the
  children of newly refined cells, did not allow different cells
  to store quadrature data of different lengths. This release lifts
  this restriction, which enables efficient quadrature data storage and data
  transfer within a triangulation containing multiple material models, each with
  their own number of local state variables and history variables.
  As an example, in a solid mechanics simulation, we can assign a hyper-elastic
  material model to a region of the mesh with no associated
  history variables, while in another part of the triangulation we incorporate an
  elasto-plastic constitutive model, requiring the storage of local plasticity
  data at quadrature points.
  During each refinement of the mesh, we can then use
  \texttt{ContinuousQuadratureDataTransfer} to transfer and interpolate
  plasticity data from parent cells to their children, ignoring the cells
  with the hyper-elastic constitutive model.
\end{itemize}
%
The changelog lists more than 240 other
features and bugfixes.

@@

Through the addition of the \texttt{Particles::ParticleHandler::insert\_global\_particles()} member function,
particles can now be inserted in parallel from a vector of points even if these points do not lie
on the subdomain from which the insertion is called. This operation requires extensive 
communication between the processes to locate   the MPI process that
owns the cell in which the particle
is located. However, it is made significantly faster  through the usage of bounding boxes
that surround the cells of each subdomain. 
This function also takes care of transferring the properties attached to the particles to their new owner.
This new capability enables particle generators that insert particles at the location
of the support points (\texttt{Particles::Generator::dof\_support\_points()}) and at the quadrature points (\texttt{Particles::Generator::quadrature\_points()}) of a
possibly non-matching triangulation. Consequently, complex particle patterns can be inserted
using unstructured grids generated outside of \dealii{}.

@@

\begin{table}
\caption{\it Supported vector lengths of the class \texttt{VectorizedArray} and
the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
\centering
\begin{tabular}{ccc}
\toprule
\textbf{double} & \textbf{float} & \textbf{ISA}\\
\midrule
VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2/AltiVec \\
VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\
VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512 \\
\bottomrule
\end{tabular}

@@

\item \texttt{step-58} is a program that solves the nonlinear
  Schr{\"o}dinger equation, which in non-dimensional form reads
  \begin{align*}
  - i \frac{\partial \psi}{\partial t}
  - \frac 12 \Delta \psi
  + V \psi
  + \kappa |\psi|^2 \psi
  &= 0,
  \end{align*}
  augmented by appropriate initial and boundary conditions and using
  an appropriate form for the potential $V=V(\mathbf x)$. The
  tutorial program focuses on two specific aspects for which this
  equation serves as an excellent test case: (i) Solving
  complex-valued problems without splitting the equation into its
  real and imaginary parts (as \texttt{step-29} does, for
  example). (ii)~Using operator splitting techniques. The equation is
  a particularly good test case for this technique because the only
  nonlinear term, $\kappa |\psi|^2 \psi$, does not contain any
  derivatives and consequently forms an ODE at each node point, to be
  solved in each time 
  step in an operator splitting scheme (for which, furthermore, there
  exists an analytic solution), whereas the remainder of the
  equation is linear and easily solved using standard finite element
  techniques.

@@

\item \texttt{step-70}
  \todo[inline]{Luca, please write.}
\end{itemize}

@@

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
%
% Uwe Koecher doesn't usually show up in the changelog, but
% we should make sure he's listed.
%
% 9.2: updated 5/11/2020 MM
\todo[inline]{TODO: remove authors of paper}
Pasquale Africa,
Ashna Aggarwal,
Giovanni Alzetta,
Mathias Anselmann,
Daniel Arndt,
Wolfgang Bangerth,
Kirana Bergstrom,
Manaswinee Bezbaruah,
Bruno Blais,
Benjamin Brands,
Yong-Yong Cai,
Fabian Castelli,
Thomas C. Clevenger,
Katherine Cosburn,
Denis Davydov,
Elias Dejene,
Stefano Dominici,
Brett Dong,
Luel Emishaw,
Marc Fehling,
Niklas Fehn,
Rebecca Fildes,
Menno Fraters,
Andres Galindo,
Daniel Garcia-Sanchez,
Rene Gassmoeller,
Melanie Gerault,
Nicola Giuliani,
Brandon Gleeson,
Anne Glerum,
Krishnakumar Gopalakrishnan,
Graham Harper,
Mohammed Hassan,
Nicole Hayes,
Bang He,
Johannes Heinz,
Timo Heister,
Luca Heltai,
Jiuhua Hu,
Lise-Marie Imbert-Gerard,
Manu Jayadharan,
Daniel Jodlbauer,
Marie Kajan,
Guido Kanschat,
Alexander Knieps,
Uwe K{\"o}cher,
Martin Kronbichler,
Paras Kumar,
Konstantin Ladutenko,
Charu Lata,
Adam Lee,
Wenyu Lei,
Matthias Maier,
Katrin Mang,
Mae Markowski,
Franco Milicchio,
Adriana Morales Miranda,
Peter Munch,
Bob Myhill,
Emily Novak,
Omotayo Omosebi,
Alexey Ozeritskiy,
Jean-Paul Pelteret,
Rebecca Pereira,
Geneva Porter,
Laura Prieto Saavedra,
Reza Rastak,
Roland Richter,
Jonathan Robey,
Irabiel Romero,
Matthew Russell,
Tonatiuh Sanchez-Vizuet,
Natasha S. Sharma,
Doug Shi-Dong,
Konrad Simon,
Stephanie Sparks,
Sebastian Stark,
Simon Sticko,
Jan Philipp Thiele,
Jihuan Tian,
Ignacio Tomas,
Sara Tro,
Bruno Turcksin,
Ferdinand Vanmaele,
Zhuoran Wang,
David Wells,
Michal Wichrowski,
Julius Witte,
Winnifried Wollner,
Ming Yang,
Mario Zepeda Aguilar,
Wenjuan Zhang,
Victor Zheng.

@@

>>>>

\hypersetup{
  pdfauthor={
     Daniel Arndt,
     Wolfgang Bangerth,
     Bruno Blais,
     Thomas C. Clevenger,
%     Denis Davydov,
    Marc Fehling,
%     Daniel Garcia-Sanchez,
    Alexander V. Grayver,
%     Graham Harper,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Peter Munch,
%     Ross Maguire Kynch,
    Matthias Maier,
    Jean-Paul Pelteret,
    Reza Rastak,
    Bruno Turcksin,
    Zhuoran Wang,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.2, 2020},
}

@@

\author[*1]{Bruno~Turcksin}
%

\author[14]{Zhuoran Wang}
\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
          CO 80523-1874, USA.
{\texttt{zhrwang@math.colostate.edu}}}

\author[15]{David Wells}
\affil[15]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

\renewcommand{\labelitemi}{--}

@@

The major changes of this release are:
%
\begin{itemize}
  \item xy \todo[inline]{Update once we have the subsections in Section 2;
          provide cross-references to each of these subsections}
        %   \item Improved support for automatic differentiation (see
        %     Section~\ref{subsec:ad}),
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item \dealii{} had decent support for solving complex-valued problems
        for a while already
        (e.g., ones in quantum mechanics -- like the equation used in the
        \texttt{step-58} tutorial program covered below -- or for
        time-harmonic problems). However, there were two
        areas in which support was missing. First, the UMFPACK direct solver
        packaged with \dealii{} did not support solving complex-valued
        linear problems. This has now been addressed: UMFPACK actually can
        solve such systems, we just needed to write the appropriate
        interfaces. Second, the \texttt{DataOut} class that is responsible
        for converting nodal data into information that can then be written
        into files for visualization did not know how to deal with vector-
        and tensor-valued fields whose components are complex numbers. An
        example for this is to solve the time-harmonic version of the
        Maxwell equations that has the electric and magnetic fields as
        solution. This, too, has been addressed in this release.
  \item The new \texttt{DiscreteTime} class provides a more
        consistent, more readable, and less error-prone approach to control
        time-stepping algorithms within time-dependent simulations.
        The interface of this class is designed to be minimal
        to enforce a number of important programming invariants, reducing
        the possibility of mistakes in the user code.
        An example is that \texttt{DiscreteTime} ensures that the final time step ends
        precisely on a predefined end time, automatically
        lengthening or shortening to final time step.
  \item A key component of \dealii{} are the \texttt{FEValues} and
        \texttt{FEFaceValues} classes that evaluate finite element functions
        at quadrature points located on cells and faces of a cell,
        respectively~\cite{BangerthHartmannKanschat2007}. This release now
        contains a class \texttt{FEInterfaceValues} that considers the
        restriction of the shape functions from \textit{both} sides of a face and
        allows evaluating jumps and averages of shape functions along this
        face. These are common components of the bilinear forms of
        discontinuous Galerkin schemes (as well as schemes for fourth-order
        equations, see the discussion of \texttt{step-47} below) and greatly
        simplify the implementation of these methods.
  \item Previously, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer}
        class, which transfers local quadrature point data onto the
        children of newly refined cells, did not allow different cells
        to store quadrature data of different lengths. This release lifts
        this restriction, which enables efficient quadrature data storage and data
        transfer within a triangulation containing multiple material models, each with
        their own number of local state variables and history variables.
        As an example, in a solid mechanics simulation, we can assign a hyper-elastic
        material model to a region of the mesh with no associated
        history variables, while in another part of the triangulation we incorporate an
        elasto-plastic constitutive model, requiring the storage of local plasticity
        data at quadrature points.
        During each refinement of the mesh, we can then use
        \texttt{ContinuousQuadratureDataTransfer} to transfer and interpolate
        plasticity data from parent cells to their children, ignoring the cells
        with the hyper-elastic constitutive model.
\end{itemize}
%
The changelog lists more than 240 other
features and bugfixes.

@@

Through the addition of the \texttt{Particles::ParticleHandler::insert\_global\_particles()} member function,
particles can now be inserted in parallel from a vector of points even if these points do not lie
on the subdomain from which the insertion is called. This operation requires extensive
communication between the processes to locate   the MPI process that
owns the cell in which the particle
is located. However, it is made significantly faster  through the usage of bounding boxes
that provide a coarse description of the geometrical shape of each subdomain.
This function also takes care of transferring the properties attached to the particles to their new owner.
This new capability enables particle generators that insert particles at the location
of the support points (\texttt{Particles::Generator::dof\_support\_points()}) and at the quadrature points (\texttt{Particles::Generator::quadrature\_points()}) of a
possibly non-matching triangulation. Consequently, complex particle patterns can be inserted
using unstructured grids generated outside of \dealii{}.

@@

\begin{table}
  \caption{\it Supported vector lengths of the class \texttt{VectorizedArray} and
    the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
  \centering
  \begin{tabular}{ccc}
    \toprule
    \textbf{double}            & \textbf{float}             & \textbf{ISA}         \\
    \midrule
    VectorizedArray<double, 1> & VectorizedArray<float, 1>  & (auto-vectorization) \\
    VectorizedArray<double, 2> & VectorizedArray<float, 4>  & SSE2/AltiVec         \\
    VectorizedArray<double, 4> & VectorizedArray<float, 8>  & AVX/AVX2             \\
    VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512              \\
    \bottomrule
  \end{tabular}

@@

  \item \texttt{step-58} is a program that solves the nonlinear
        Schr{\"o}dinger equation, which in non-dimensional form reads
        \begin{align*}
          - i \frac{\partial \psi}{\partial t}
          - \frac 12 \Delta \psi
          + V \psi
          + \kappa |\psi|^2 \psi
           & = 0,
        \end{align*}
        augmented by appropriate initial and boundary conditions and using
        an appropriate form for the potential $V=V(\mathbf x)$. The
        tutorial program focuses on two specific aspects for which this
        equation serves as an excellent test case: (i) Solving
        complex-valued problems without splitting the equation into its
        real and imaginary parts (as \texttt{step-29} does, for
        example). (ii)~Using operator splitting techniques. The equation is
        a particularly good test case for this technique because the only
        nonlinear term, $\kappa |\psi|^2 \psi$, does not contain any
        derivatives and consequently forms an ODE at each node point, to be
        solved in each time
        step in an operator splitting scheme (for which, furthermore, there
        exists an analytic solution), whereas the remainder of the
        equation is linear and easily solved using standard finite element
        techniques.

@@

  \item \texttt{step-70} solves a fluid structure interaction problem on
        non-matching parallel distributed triangulations, showing the usage of
        the \texttt{particles::ParticleHandler} class for two different tasks:
        i) to track the position of quadrature points of a non-matching
        solid grid immersed in a fluid grid, and ii) to track the position of a
        collection of massless tracers.

        In \texttt{step-70} we consider a mixing problem in the laminar flow
        regime. Such problems occur in a wide range of applications ranging
        from chemical engineering to power generation (e.g. turbomachinery).
        Mixing problems are particularly hard to solve numerically, because
        they often involve a container (with fixed boundaries, and possibly
        complex geometries such as baffles), and one (or more) immersed and
        rotating impellers, making it difficult to use Arbitrary Lagrangian
        Eulerian formulations, where the fluid domain – along with the mesh! –
        is smoothly deformed to follow the deformations of the immersed solid.
        We show how to solve such problems using a penalization technique
        where the flow problem is solved in the union of the fluid and solid
        domain, and the movement of the solid is imposed weakly using a
        Nitsche-like penalization.

        The \texttt{particles::ParticleHandler} class is used in this context
        to allow integration of the fluid basis functions on the solid domain
        (which is immersed and non-matching w.r.t. the fluid grid). This is
        achieved by attaching the information required to perform the
        integration on the solid grid, to the property field of the particles
        associated with the quadrature points of the solid, and advecting those
        particles according to the solid velocity. In this program, we assign
        the solid velocity externally, and we do not solve for it, even though
        the program can be easily extended to allow for the solution of
        coupled elasticity equations on the solid domain.
\end{itemize}

@@

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
%
% Uwe Koecher doesn't usually show up in the changelog, but
% we should make sure he's listed.
%
% 9.2: updated 5/11/2020 MM
Pasquale Africa,
Ashna Aggarwal,
Giovanni Alzetta,
Mathias Anselmann,
Kirana Bergstrom,
Manaswinee Bezbaruah,
Benjamin Brands,
Yong-Yong Cai,
Fabian Castelli,
Katherine Cosburn,
Denis Davydov,
Elias Dejene,
Stefano Dominici,
Brett Dong,
Luel Emishaw,
Niklas Fehn,
Rebecca Fildes,
Menno Fraters,
Andres Galindo,
Daniel Garcia-Sanchez,
Rene Gassmoeller,
Melanie Gerault,
Nicola Giuliani,
Brandon Gleeson,
Anne Glerum,
Krishnakumar Gopalakrishnan,
Graham Harper,
Mohammed Hassan,
Nicole Hayes,
Bang He,
Johannes Heinz,
Jiuhua Hu,
Lise-Marie Imbert-Gerard,
Manu Jayadharan,
Daniel Jodlbauer,
Marie Kajan,
Guido Kanschat,
Alexander Knieps,
Uwe K{\"o}cher,
Paras Kumar,
Konstantin Ladutenko,
Charu Lata,
Adam Lee,
Wenyu Lei,
Katrin Mang,
Mae Markowski,
Franco Milicchio,
Adriana Morales Miranda,
Bob Myhill,
Emily Novak,
Omotayo Omosebi,
Alexey Ozeritskiy,
Rebecca Pereira,
Geneva Porter,
Laura Prieto Saavedra,
Roland Richter,
Jonathan Robey,
Irabiel Romero,
Matthew Russell,
Tonatiuh Sanchez-Vizuet,
Natasha S. Sharma,
Doug Shi-Dong,
Konrad Simon,
Stephanie Sparks,
Sebastian Stark,
Simon Sticko,
Jan Philipp Thiele,
Jihuan Tian,
Sara Tro,
Ferdinand Vanmaele,
Michal Wichrowski,
Julius Witte,
Winnifried Wollner,
Ming Yang,
Mario Zepeda Aguilar,
Wenjuan Zhang,
Victor Zheng.

@@

L.~Heltai was partially supported by the Italian Ministry of Instruction,
University and Research (MIUR), under the 2017 PRIN project NA-FROM-PDEs MIUR
PE1, ``Numerical Analysis for Full and Reduced Order Methods for the efficient
and accurate solution of complex systems governed by Partial Differential
Equations''.
